### 2024-11-29 paper/writeup.tex
Tweaks

\documentclass[10pt,twoside,lineno]{gsajnl}
\articletype{inv} % article type

% JK Had to change the xcolor import here and turn off
% the OliveGreen bit below to get this to compile
% \usepackage[dvipsnames]{xcolor}
\usepackage[]{xcolor}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage[linesnumbered,ruled,vlined,algo2e]{algorithm2e}
\usepackage{tikz}
\usetikzlibrary{arrows, snakes,backgrounds}
\tikzstyle{place}=[circle,draw=black,thick, inner sep=0pt, minimum size = 5mm]
\usepackage{hyperref}
\hypersetup{
	pagebackref=true,
	colorlinks = true,
}
	% allcolors = OliveGreen}

\SetKwInput{input}{Input}
\SetKwInput{output}{Output}

\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\T}{\mathbb{T}}
\newcommand{\ind}{\mathbf{1}}
\newcommand{\tn}{\textnormal}
\newcommand{\ov}{\overline}
\newcommand{\tskit}{\texttt{tskit}}
\newcommand{\tsinfer}{\texttt{tsinfer}}
\newcommand{\msprime}{\texttt{msprime}}
\newcommand{\argmax}{\operatorname{argmax}}
\newcommand{\dis}{\operatorname{dis}}
\newcommand{\similarity}{\operatorname{sim}}
\newcommand{\comment}[1]{{\color{violet} \it #1}}
\newcommand{\pluseq}{\mathrel{{+}{=}}}

\newcommand{\algorithmref}[2][]{%
	\hyperref[{#2}]{%
		Algorithm~\ref*{#2}%
		\ifx\\#1\\%
		\else
		\,#1%
		\fi
	}%
}

\newcommand*{\figref}[2][]{%
	\hyperref[{#2}]{%
		Figure~\ref*{#2}%
		\ifx\\#1\\%
		\else
		\,#1%
		\fi
	}%
}

\newcommand*{\tabref}[2][]{%
	\hyperref[{#2}]{%
		Table~\ref*{#2}%
		\ifx\\#1\\%
		\else
		\,#1%
		\fi
	}%
}
\newcommand*{\eqnref}[2][]{%
	\hyperref[{#2}]{%
		Equation~\ref*{#2}%
		\ifx\\#1\\%
		\else
		\,#1%
		\fi
	}%
}

\title{
    A forest is more than its trees:
    haplotypes and inferred ARGs
}

% Information in recombination junctions can both compress and improve inferred tree sequences
% Non-coalescing regions of dark matter in ARGs
% Can we find them? Are they useful in inference, dating? 
% Look, we can find them and they seem at least useful for compression.
% Also, here's a way of measuring agreement that takes this sort of thing into account,
%  which extends R-F to measure haplotypes

\author[$\dagger$]{Halley Fritze}
\author[$\ast$]{Nathaniel Pope}
\author[$\ddagger$]{Jerome Kelleher}
\author[$\ast$,$\dagger$,S]{Peter Ralph}

\affil[$\ast$]{Institute of Evolution and Ecology and Department of Biology, University of Oregon, Eugene, Oregon}
\affil[$\dagger$]{Department of Mathematics, University of Oregon, Eugene, Oregon}
\affil[$\ddagger$]{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, University of Oxford}
\affil[$\S$]{Department of Data Science, University of Oregon, Eugene, Oregon}

\keywords{genealogy, tree sequence, haplotypes, ancestral recombination graph}

\runningtitle{Haplotypes in ARGs}
\runningauthor{Fritze \textit{et al.}}

%%%%%%%%%%
\begin{abstract}
    Foreshadowing haplotype-based methods of the genomics era,
    it is an old observation that the ``junction'' between two distinct haplotypes
    produced by recombination is inherited as a Mendelian marker.
    In this paper, we describe how this recombination-mediated information
    can in many cases be recovered from inference based solely on polymorphic markers.
    In a genealogical context, this information reflects the persistence
    of ancestral haplotypes across marginal genealogical trees in which they
    do not represent coalescences.
    We show how these non-coalescing haplotypes (``unary regions'') may be
    inserted into the ancestral recombination graph (ARG), a compact
    but information-rich data structure describing the genealogical
    relationships among recombinant sequences.
    The resulting ARGs are smaller, faster to compute with,
    and the additional ancestral information that is inserted
    is nearly always correct where the initial ARG is correct.
    We provide efficient algorithms to infer unary regions within existing ARGs,
    and explore some consequences for ARGs inferred from real data.
    To do this, we introduce new metrics of agreement and disagreement between inferred ARGs
    that, unlike previous methods, consider the ARG as describing relationships between haplotypes
    rather than just a collection of trees.
\end{abstract}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\marginmark
\firstpagefootnote

\correspondingauthoraffiliation{1}{Corresponding author: {plr@uoregon.edu}}
\vspace{-33pt}% Only used for adjusting extra space in the left column of the first page

%%% OUTLINE
% Intro: haplotypes, unary bits of non-coalescing nodes, ARGS, etcetera
%    explain what tsinfer does to create unary regions
%    statement of problem
%    Yan: are breakpoints in true->simplify->extend the same as in original?
%   Fig 1: conceptual figure
%    Also IBD gets screwed up
% 
% Methods:
%   edge extend algorithm
%   discrepancy funciton algorithm
%   Fig 2: Conceptual figure for discrepancy
% 
% Results:
%   Fig 3:
%     (a) reduction in number of edges and (b) speed change
%
%   Fig 4:
%     Histograms of (a) total span added, (b) percent incorrect
%
%   Fig 5 and maybe 6: how it interacts with tsinfer
%     (a) summarize total matching and unmatching span (or maybe percent matching?)
%           across reps
%     (b) discrepancy per node against depth or time or number of subtended samples
%        describe percent span matched against true span or depth or something
%
%   Fig 7(?): compare IBD stats before/after
% 
% Supp:
%   S1: runtime
%%%

% Name ideas:
% 
% extend edges
% bundle edges
% inflate edges
% bundle lines of descent
% longer ancestral haplotypes
% inflated ancestors
% compress paths
% optimizing edge tables
% reduce number of ancestors
% reduce ancestral paths

%%%% COLOR SCHEME FOR FIGURES %%%%
% colors = {'blue': 'rgb(46,37,133)',
%	'red': 'rgb(194,106,119)',
%	'lgreen': 'rgb(93,168,153)',
%	'gold': 'rgb(220,205,125)',
%	'green': 'rgb(51, 117,56)',
%	'lblue': 'rgb(148,203,236)',
%	'magenta': 'rgb(159,74,150)',
%	'wine': 'rgb(126,041,084)', 
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

% PETER
% * what's a tree sequence
% * why is a tree sequence (motivation)

Ancestral recombination graphs (ARGs) describe how a set of sampled sequences 
are related to each other at each position of the genome in a recombining
species~\citep{brandt2024promise,lewanski2024introduction,nielsen2024inference,
wong2024general}, and there has been significant recent progress 
on inference through a range of approaches~\citep{rasmussen2014genome,
speidel2019method,kelleher2019inferring,zhang2023biobankscale,deng2024robust}.
One way of viewing ARGs is as a sequence of marginal trees,
i.e., the genealogical trees that describe how each portion of the genome
was inherited by the focal genomes.
This is reflected in methodology of some ARG inference methods
and in metrics used to assess inference accuracy,
as well as in basic terminology.
For instance, 
the ``succinct tree sequence'',
introduced by \citet{kelleher2016efficient},
is a common format for describing these inferred ARGs,
and is seeing wide use thanks in part to its efficiency and accompanying reliable toolkit,
\tskit{} \citep{tskit,ralph2020efficiently}.

However, an ARG is emphatically not merely a sequence of trees:
viewed another way, it describes inheritance relationships between ancestral haplotypes.
These two points of view are related because a single haplotype
may extend over many marginal trees;
in other words, the internal nodes in the trees are labeled, and many of these labels
are shared between adjacent trees~\citep{wong2024general}.

Another reason we tend to focus on the trees is that
much of our intuition about inference of relationships from genomic data
comes from phylogenetics.
Indeed, all methods might very roughly be summarized as
``more similar sequences are more closely related''.
For instance, two sequences that share a derived mutation
are probably more closely related over some span of genome surrounding the location where the mutation occurs.
It has long been observed 
that not only mutations
but also the ``junctions'' between distinct haplotypes,
if they could be somehow identified,
would be inherited as Mendelian markers
\citep{fisher1954fuller,chapman2003model}.
In more modern terminology, 
even in the absence of new mutations,
recombination between distinct haplotypes can create a novel haplotype
whose relationships and origination time could be inferred.

Haplotype identity has been largely overlooked in the literature on ARG inference --
to our knowledge, all metrics that have been used so far to measure accuracy of inferred ARGs
depend only on the sequence of marginal trees,
not on how ancestral haplotypes span across these trees.
For instance, \citet{kelleher2019inferring} and \citet{zhang2023biobankscale}
compared true and inferred ARGs
using average Robinson-Foulds~\citep{robinson1981comparison} and 
Kendall-Colijn~\citep{kendall2016mapping} distances between trees
across a regular sequence of genomic positions,
using sampled genotypes as labels,
while \citet{brandt2022evaluation} compared times to most recent common ancestor
between pairs of sampled genomes.
Neither is affected by shared haplotype structure --
two ARGs could be identical by either measure
but imply completely different patterns of haplotype sharing and inheritance.
% (as defined in the context of the ARG, \citep{XXX}).
\citet{deng2021distribution} evaluated agreement of distributions of
distances along the genome between tree topology changes,
and \citet{zhang2023biobankscale} defined a generalization of Robinson-Foulds distance
that is the total variation distance between the induced distribution on genotypes;
neither of which measure the sharing of haplotypes between adjacent trees.
However, the additional information provided by haplotype structure can be important:
for instance, haplotypes that extend over many marginal genealogies
``tie together'' those genealogies, allowing estimates of times of particular ancestors
to be informed by larger portions of the genome on which there are many genealogies.

In this paper, we study various aspects of haplotype identity in ARGs.
First, we describe a deterministic algorithm that
extends the genomic region spanned by ancestral haplotypes
using the principle that intermediate nodes in inheritance paths
should remain unchanged when possible.
These extended portions of ancestral haplotypes manifest as unary nodes in the marginal trees.
To quantify how accurate the new information is,
we define and describe how to compute new measures of (dis-)agreement between ARGs
that are motivated by the Robinson-Foulds distance between trees
but account for haplotype identity.
These measures show that the vast majority of these extended haplotypes are correct if the trees are correct,
and that substantial information about haplotypes is contained in these nodes
in inferred trees as well.

\subsection{Motivation and statement of problem}

% PETER
% * minimize number of edges
% * gives extra info about shared haplotypes, reduces number of ancestral paths

Consider the (small portion) of a hypothetical ARG in \figref[A]{fig:conceptual}.
On the first portion of the genome (left-hand tree), the sample nodes (labeled 0, 1, and 2)
coalesce into a small subtree: 1 and 2 find a common ancestor in ancestral node 3,
which finds a common ancestor with node 0 in ancestral node 4.
On the next portion of the genome (right-hand tree), sample node 2 has a different ancestor.
This seems reasonable, and a method that infers trees separately on each portion of the genome
could not be expected to produce anything different.
However, the example becomes more complicated once we consider what these marginal genealogies imply about haplotype inheritance.
\figref[B]{fig:conceptual} shows the implied inheritance of haplotypes,
with the haplotypes carried by 4 to the left and right of the recombination breakpoint labeled $L$ and $R$.
Here, sample node 2 has inherited the chunk of haplotype labeled $L$ from ancestral node 4 via 3,
and the haplotype to the right of this from some other node (and so doesn't carry haplotype $R$).
On the other hand, sample node 1 has inherited \emph{both} haplotypes $L$ and $R$
from ancestral node 4, but the trees imply that only haplotype $L$ is inherited via ancestral node 3.
This implies -- if taken literally -- that there must have been a recombination event
at some point between node 1 and node 4 that separated the $L$ and $R$ haplotypes,
and then these two ancestral (and nonoverlapping) haplotypes coalesced together in ancestral node 4.
Although this is possible, it seems unlikely --
a more parsimonious explanation is depicted in \figref[C]{fig:conceptual},
in which sample node 1 inherits the entire $LR$ haplotype from ancestral node 4 through node 3
(and there is a recombination somewhere between node 3 and node 2).
This implies that ancestral node 3 inherits from node 4 on the right-hand tree as well,
which is depicted in \figref[D]{fig:conceptual} --
and so node 3 has become unary in this tree.
Note that the more parsimonious ARG also includes fewer edges:
the three distinct edges $4 \to 3$, $3 \to 1$, and $4 \to 1$ in \figref[B]{fig:conceptual}
have been reduced to the two edges
$4 \to 3$ and $3 \to 1$ in \figref[D]{fig:conceptual}.

So, given the ARG shown in \figref[A\&B]{fig:conceptual},
it should be possible to extend the ancestral haplotype represented by node 3
to obtain the ARG shown in \figref[C\&D]{fig:conceptual},
thus adding additional information to the ARG.
This might be surprising,
as intuition from phylogenetics suggests we can only infer
information about the branching points in the tree, not intermediate (unary) nodes.
The goal of this paper is to answer:
How can we do this, and how accurate is the resulting inference?

\begin{figure}
    \begin{center}
    \includegraphics[width=0.9\textwidth]{conceptual_figure.pdf}
    \end{center}
    \caption{
        A simple example showing the basic idea
        (described in more detail in the text):
        \textbf{(A)} a small portion of an ARG without unary nodes;
        \textbf{(B)} the implied inheritance pattern of the two portions of the haplotype carried by ancestral node 4,
        labeled $L$ and $R$;
        \textbf{(C)} marginal trees with a unary node added,
        which produces \textbf{(D)} a more parsimonious haplotype inheritance pattern
        (that also includes fewer edges).
        \label{fig:conceptual}
    }
\end{figure}

\section{Methods}

\subsection{Notation and terminology}

We work with the \emph{succinct tree sequence} representation of ARGs (henceforth, ``tree sequence''),
to take advantage of the extensive, stable, and well-documented set of open-source tools
available in \tskit{} \citep{tskit}, 
and our terminology and notation follows \citet{ralph2020efficiently}.
For our purposes here,
a tree sequence $\T = (N, E)$ contains a set of \emph{nodes} $N$ 
which represent ancestral segments of genome,
and \emph{edges} $E$ which represent relationships between nodes over different regions of the genome.
Each node $n \in N$ has a \emph{time} $t_n$,
which is the amount of time in the past that the individual who carried that segment of genome lived.
Some nodes are \emph{samples}, meaning that they represent genome sequences available as data.
% Typically, samples are exclusively leaves of $\T$, however it is not the case that all samples are leaves.
Each edge $e \in E$ describes inheritance between a parent node $p_e$ and child node $c_e$,
over a segment of genome $[\ell_e, r_e)$.
(The tree sequence data structure typically includes substantially more information, including genotypes, that we will not use in this paper.)
Suppose that the unique elements of the set of left and right edge endpoints
are $0 = a_0 < a_1 < \cdots < a_{n} = L$, where $L$ is the length of the genome.
Using this information, one can construct the sequence of 
trees $\left(T_1,...,T_{n}\right)$ that describe how the nodes are related to each other along the genome:
each $T_k$ is a tree whose nodes are in $N$
and that describes relationships on the half-open interval $[a_{k-1}, a_k)$.
Nodes represent (portions of) ancestral haplotypes, and so
we will use the terms interchangeably.
Not all nodes appear in each tree,
and we say $n \in T_k$ for a node $n$ if the tree $T_k$ describes at least one parent-child relationship
for node $n$.

\subsection{An algorithm to extend haplotypes}

% Description of algorithm
Given a tree sequence, our goal is to
identify areas of implied inheritance of haplotypes.
Generalizing from \figref{fig:extending_diagram},
we do this by identifying paths of inheritance that are shared across a sequence of marginal trees
but for which some of the intermediate nodes are missing.
Concretely, suppose that 
if in tree $T_k$ there is a chain of inheritance
$p \to u_1 \to \cdots \to u_m \to c$
(where $a \to b$ denotes a parent-child relationship)
and in tree $T_{k+1}$ there is a chain of inheritance
$p \to v_1 \to \cdots \to v_n \to c$,
where $\{u_i\}_{i=1}^m$ and $\{v_j\}_{j=1}^n$ are disjoint.
This situation implies that $c$ inherited from $p$ over the entire interval $[a_{k-1}, a_{k+1})$,
so it seems reasonable to assume that $c$ has inherited from $p$ \emph{along the same path} for that entire interval.
In other words, the intermediate nodes $\{u_i\}$ should also lie on the path from $c$ to $p$ in tree $T_{k+1}$,
and conversely the nodes $\{v_j\}$ should lie on that path in tree $T_k$.
Of course, this does not always make sense --
for instance, if $u_i$ is already represented somewhere else in $T_{k+1}$,
or if $t_{u_i} = t_{v_j}$, for some $i$ and $j$.
So, we restrict our attention to pairs of such paths in adjacent trees
for which
$u_i \notin T_{k+1}$ for all $1 \le i \le m$,
$v_j \notin T_k$ for all $1 \le j \le n$,
and the times of the nodes $\{u_i\}$ and $\{v_j\}$ are unique.
Call a pair of such paths \emph{mergeable}.
So, the goal of our algorithm is to iterate over trees,
identify mergeable pairs of paths,
and then extend the nodes $\{u_i\}$ to $T_{k+1}$.
(We also extend $\{v_j\}$ to $T_k$, but on a backwards pass.)

\begin{figure}[!ht]
\begin{center}
	\includegraphics[width=5in]{edge_extend_method.pdf}
\end{center}
\caption{A visualization of the \textit{extend haplotypes} method.
    In both trees $T_1$ and $T_2$,
    node 0 inherits from node 6, the root:
    $T_1$ contains the path $6\to 4\to 0$ while $T_2$ has path $6\to 5\to 0$.
    The intermediate nodes $4$ and $5$ do not appear in $T_2$ and $T_1$ respectively, and so the paths are \textit{mergeable}.
    The ``extend haplotypes'' method joins these two paths,
    inserting the merged path $6\to 5\to 4\to 0$ into both $T_1$ and $T_2$.
    % This new path extends the edge $4\to 0$ to $T_2$ as well as constructs a new edge $5\to 4$ in the tree sequence.
    \label{fig:extending_diagram}
}
\end{figure}

% JK note: earlier version said "relatively" - relative to what?
An efficient algorithm to do this is described in \algorithmref{alg:extend}, 
and implemented in \tskit{} as \texttt{extend\_haplotypes}.
The algorithm considers each tree transition from $T_k$ to $T_{k+1}$ in turn, updating its internal state
(which includes possibly modifying $T_k$) as it goes.
Suppose we are at the transition from tree $T_k$ to tree $T_{k+1}$,
which is done by first removing a set of edges $O$
and then adding another set of edges $I$.
$O$ defines a sub-forest $F_O$ of $T_k$,
and $I$ defines a sub-forest $F_I$ of $T_{k+1}$.
The key step in the algorithm is to determine whether the pair of paths
that terminate in a given node in two adjacent trees are mergeable.
The algorithm we use to do this is given as \algorithmref{alg:mergeable},
and works as follows.
If a pair of paths is mergeable,
then the edges of the two paths must lie in $O$ and $I$, respectively.
Suppose an edge in $O$ has child $c$.
To see if $c$ is the base of a pair of mergeable paths,
the algorithm traverses up from $c$ in both $F_O$ and $F_I$;
terminating if a node in the other tree is found
(i.e., if the node traversed in $F_O$ is in $T_{k+1}$ or if the node traversed in $F_I$ is in $T_k$)
or if a pair of traversed nodes have the same time.
If these two traversals end in the same node $p$, the paths are mergeable.
Iterating over all edges in $O$ will thus find all mergeable pairs of paths.
There is often more than one pair of mergeable paths in a tree transition;
so, the algorithm merges pairs of mergeable paths,
starting with pairs that add the smallest number of new edges,
until no more are found.

\algorithmref{alg:extend} simplifies the algorithm in \tskit{}
in several ways for the sake of clarity --
for instance, the bookkeeping required to keep track of $T_k$ and $T_{k+1}$ is omitted.
Furthermore, as described the algorithm does one left-to-right pass over the tree sequence;
in practice we do repeated passes in both directions until no changes can be made.
The main step that is omitted is a description of the \texttt{merge} operation,
which performs the actual extending of haplotypes.
This algorithm is essentially the same as \texttt{mergeable} in \algorithmref{alg:mergeable},
except with additional bookkeeping.
Roughly speaking, the algorithm
traverses up from the shared base node $c$,
doing the appropriate operations to insert the nodes along the path found in $T_k$
into the path in $T_{k+1}$.
To do this, some edges that end at $a_k$ will be extended to end at $a_{k+1}$;
some edges that begin at $a_k$ will be postponed to begin at $a_{k+1}$,
and some entirely new edges may be added, as in \figref{fig:extending_diagram}.
Furthermore, the trees $T_k$ and $T_{k+1}$ (and corresponding forests $F_O$ and $F_I$)
need to be updated.

\begin{algorithm2e}[!ht]
	\SetStartEndCondition{ }{}{}%
	\SetKwProg{Fn}{def}{\string:}{}
	\SetKwFunction{Range}{range}%%
	\SetKw{KwTo}{in}
	\SetKwFor{For}{for}{\string:}{}%
	\SetKwIF{If}{ElseIf}{Else}{if}{:}{elif}{else:}{}%
	\SetKwFor{While}{while}{:}{fintq}%
	\SetKw{Break}{break}
	\newcommand\forcond{$i$ \KwTo\Range{$n$}}
	\AlgoDontDisplayBlockMarkers\SetAlgoNoEnd\SetAlgoNoLine%
    \caption{
        Given a node $c$, trees $T_O$ and $T_I$,
        and sub-forests $F_O$ and $F_I$
        such that removing $F_O$ and adding $F_I$ turns $T_O$ into $T_I$,
        check to see if the paths upwards from $c$ in $T_O$ and $T_I$ are mergeable.
        If the paths are mergeable then this returns the number of new edges
        that would be added by extending the path from $T_O$ to $T_I$;
        otherwise, this returns $\infty$.
        Let $P_O[n]$ and $P_I[n]$ be the parents of node $n$
        in the set of edges to be removed and added, respectively
        (i.e., in $F_O$ and $F_I$).
        The variable $m_e$ will record the number of new edges to be added,
        and $m$ will record the number of extended haplotypes.
    }
	\label{alg:mergeable}
	\SetKwFunction{mergeable}{Mergeable}
	\DontPrintSemicolon
    \Fn{\mergeable($c$, $T_O$, $T_I$)}{
        \text{Let } $p_i = P_I[c]$, 
        $t_i = t[p_i]$,
        $p_o = P_O[c]$,
        $t_o = t[p_o]$,
        \text{ and }$m_e = m = 0$.\;
        \While{\tn{True}}{
            \text{Set } $y_i = (p_i \neq NULL)
                \& (p_i \notin T_O)
                \& (t_i < t_o)$ \;
            \text{and } $y_o = (p_o \neq NULL)
                \& (p_o \notin T_I)
                \& (t_o < t_i)$ \;
            \If{not ($y_i$ or $y_o$)}{ \Break\; }
            \eIf{$y_i$}{
                \If{$P_I[c] \neq p_i$ and $P_O[c] \neq p_i$}{
                    \text{Set } $m_e = m_e + 1$.\;
                }
                \text{Set } $c = p_i$,
                $p_i = P_I[p_i]$, \text{ and}\;
                $t_i = \text{ if }(p_i = NULL)\text{ then }\infty\text{ else }t[p_i]$ .\; 
            }{
                \If{$P_I[c] \neq p_o$ and $P_O[c] \neq p_o$}{
                    \text{Set } $m_e = m_e + 1$.\;
                }
                \text{Set }$c = p_o$,
                $p_o = P_O[p_o]$, \;
                $t_o = \text{ if }(p_o = NULL)\text{ then }\infty\text{ else }t[p_o]$ ,\;
                \text{and } $m = m + 1$ .\;
            }
        }
        \If{$m = 0$ or $p_i \neq p_o$ or $p_i = NULL$}{
            \text{Set } $m_e = \infty$. \;
        }
        \KwRet{$m_e$}\;
    }
\end{algorithm2e}

\begin{algorithm2e}[!ht]
	\SetStartEndCondition{ }{}{}%
	\SetKwProg{Fn}{def}{\string:}{}
	\SetKwFunction{Range}{range}%%
	\SetKw{KwTo}{in}
	\SetKwFor{For}{for}{\string:}{}%
	\SetKwIF{If}{ElseIf}{Else}{if}{:}{elif}{else:}{}%
	\SetKwFor{While}{while}{:}{fintq}%
	\SetKw{Break}{break}
	\newcommand\forcond{$i$ \KwTo\Range{$n$}}
	\AlgoDontDisplayBlockMarkers\SetAlgoNoEnd\SetAlgoNoLine%
    \caption{Extend haplotypes.}
	\label{alg:extend}
	\SetKwFunction{mergeable}{Mergeable}
	\SetKwFunction{merge}{Merge}
	\SetKwFunction{extend}{ExtendHaplotypes}
	\DontPrintSemicolon
    \Fn{\extend($\T$)}{
        \For{$k$ in $1 \ldots N$}{
            \text{Set } $M = 0$ \text{ and }
            $M' = \infty$.\;
            \While{$M < \infty$}{
                \For{$e \in O_k$}{
                    \text{Set } $m$ = \mergeable($c_e$, $T_k$, $T_{k+1}$).\;
                    \eIf{$m < M$}{
                        \merge($c_e$, $T_k$, $T_{k+1}$)\;
                    }{
                        \text{Set } $M' = \min(m, M)$. \;
                    }
                }
                \text{Set } $M = M'$ \text{ and }
                $M' = \infty$ .\;
            }
        }
    }
\end{algorithm2e}

Because the algorithm needs to take multiple passes over the tree sequence
in each direction,
an important practical question for this algorithm is:
how many passes do we need to do?
The algorithm is monotone (spans of ancestral nodes only increase),
so it is guaranteed to terminate in a finite number of passes,
but it is also not hard to construct pathological cases that require an arbitrary number of passes.
However, experimentation suggests that 
in practice at most five iterations are needed before the algorithm terminates.
Indeed, for even large sequences \tabref{tab:edge-counts}
shows that 99\% of all changes to an ARG occur after the first iteration,
with the algorithm always completing after four iterations.

\subsection{Dissimilarity between ARGs}
% * How to measure agreement that includes haplotypes
% * Definition and Algorithm
% * Supp fig: runtime ~ # trees, samples
% * fig: how add edges reduces discrepancy (RESULTS)
%       - compare to tsinfer
% * Show: how add edges reduces discrepancy (RESULTS)

If we begin with a tree sequence containing unary nodes,
it is straightforward to remove the portions of each node's span on which it is unary,
apply \algorithmref{alg:extend},
and quantify how much node span was correctly or incorrectly added.
However, we are also interested in whether \algorithmref{alg:extend}
improves \emph{inferred} ARG.
Since we are not aware of any current methods for measuring (dis)agreement between ARGs
that take into account haplotype identity,
we define a measure of \emph{dissimilarity} to quantify this.
The method is implemented in the \texttt{tscompare} package.

It is helpful to first describe what we compute in the simple case.
We will first simulate tree sequences in such a way that
the unary regions of coalescent haplotypes are retained.
Then, each node is present in both tree sequences,
and we can quantify, for each node, how much of their span is correct or incorrect
by comparing to the original, true tree sequence.

Now suppose that instead of comparing two tree sequences with the same set of nodes,
we wish to compare two tree sequences for which we know the sample nodes are the same
but are otherwise unclear as to the equivalency of nodes across sequences.
(For instance, with a simulated tree sequence and one inferred from its genotypes;
nodes in the former represent actual ancestral haplotypes, and in the latter
represent hypothetical ancestors which may or may not resemble the truth.)
Call the two tree sequences $\T_1$ and $\T_2$, which should have the same genome length
and the same set of sample nodes.
We would like to measure (a) how much of $\T_1$ is found in $\T_2$;
(b) how much of $\T_1$ is \emph{not} found in $\T_2$; and
(c) how much of $\T_2$ is not found in $\T_1$.
(Think of these three quantities as the size of the intersection
and two differences between the tree sequences,
thought of vaguely as sets.)
Roughly speaking, we first identify matching nodes
as those whose sets of descendant samples agree for the largest span along the genome,
and then compute for how much of their spans do their descendant samples agree (or not).

Concretely, our method can be applied to any two tree sequences,
as illustrated in \figref{fig:conceptual_discrepancy}.
However, to simplify notation suppose that the two tree sequences have the same set of breakpoints
between trees,
so that $T_1^{(1)}, \ldots, T_N^{(1)}$ are the trees in $\T_1$
and $T_1^{(2)}, \ldots, T_N^{(2)}$ are the trees in $\T_2$.
For a node $n$ and tree $T$
let $S(T, n)$ denote the set of samples that inherit from $n$ in $T$,
and for a pair of nodes $n_1$ and $n_2$ with $n_1$ in $\T_1$ and $n_2$ in $\T_2$,
define
\begin{align*}
	\mathcal{M}(n_1, n_2) = \left\{ k : S\left(T^{(1)}_k,\ n_1\right) = S\left(T^{(2)}_k,\ n_2\right) \right\},
\end{align*}
to be the indices of all trees where $n_1$ and $n_2$ are ancestral to the same sample set in both ARGs.
Further, we define
\begin{align*}
    m(n_1, n_2)
    =
    \sum_{k\in \mathcal{M}(n_1, n_2)} (a_k - a_{k-1}) ,  
\end{align*}
which is the total span over which the samples below $n_1$ in $\T_1$
matches the samples below $n_2$ in $\T_2$.
The \emph{similarity} between $\T_1$ and $\T_2$ is then defined to be
\begin{align*}
    \similarity(\T_1, \T_2)
    =
    \max_{\beta:N_1 \to N_2} \sum_{n \in N_1} m(n, \beta(n)) ,
\end{align*}
where the maximum is over all mappings $\beta$ of nodes in $\T_1$ to nodes in $\T_2$, and we require that samples in $\T_1$ are mapped to samples in $\T_2$.
(Note that multiple nodes in $\T_1$ may be mapped to the same node in $\T_2$,
and that there some nodes in $\T_2$ may not be mapped to by any nodes in $\T_1$.)
Since the maximum is independent over nodes, we may define for each node $n_1 \in \T_1$
its' \emph{best matching node} in $\T_2$ as
\begin{align*}
    \alpha(n_1) = \argmax_{n_2 \in N_2} m(n_1, n_2) ,
\end{align*}
so that
\begin{align*}
    \similarity(\T_1, \T_2)
    =
    \sum_{n \in N_1} m(n, \alpha(n)) .
\end{align*}
If the best-matching node is not unique, we define $\alpha(n_1)$ to be the node in $T_2$
out of those maximizing $m(n_1, n_2)$ that minimizes $|t^{(1)}_{n_1} - t^{(2)}_{n_2}|$
(and if \emph{this} is not unique, we pick an arbitrary one) --
however, this potential ambiguity does not affect the definition of $\similarity(\T_1, \T_2)$.
Let $\|\T_1\|=\sum_{n\in N_1}s(\T_1,n)$ be the total sum of nodes spans for nodes in $\T_1$,
and let $s(\T,n)$ denote
the total span that node $n$ is present in the marginal trees,
\begin{align*}
    s(\T, n) = \sum_{k=1}^N (a_k - a_{k-1}) \ind_{n \in T_k} \quad ,
\end{align*}
where $\ind_{n \in T_k}$ is an indicator (i.e., it is 1 if $n \in T_k$ and 0 otherwise).
We then define the \emph{dissimilarity} of $\T_1$ from $\T_2$ by
\begin{align*}
    \dis(\T_1, \T_2)
    =
    \sum_{n \in N_1} (s(\T_1, n) - m(n, \alpha(n))) 
    = 
    \Vert \T_1\Vert -\similarity(\T_1,\T_2),
\end{align*}
which is the total span for all nodes in $\T_1$
over which their descendant samples do \emph{not} match those of their best match in $\T_2$.

A common measure of disagreement between ARGs, 
first proposed by~\cite{kuhner2015assessing}, 
is to use a weighted average Robinson-Foulds (RF) distance.
This could be computed in a very similar way:
instead of $m(n, \alpha(n))$ define
\begin{align*}
    \mathcal{M}'(\T_1, n; \T_2) = \left\{ k : \exists n_2  \;\text{for which}\; S\left(T^{(1)}_k,\ n\right) = S\left(T^{(2)}_k,\ n_2\right) \right\},
\end{align*}
the indices of all trees on which there is \emph{some} node in $\T_2$
whose set of descendant samples matches those of $n$, and
\begin{align*}
    m'(n, \T_2) = \sum_{k \in \mathcal{M}'(\T_1, n; \T_2)}^N (a_k - a_{k-1})
\end{align*}
the total span over which $n$ finds a match.
Then the average RF distance (averaged over locations in the genome)
is
\begin{align*}
    \frac{1}{L} \left( \sum_{n_1 \in N_1} m'(n_1, \T_2)  + \sum_{n_2 \in N_2} m'(n_2, \T_1) \right).
\end{align*}
In other words, we require a node in $\T_1$ to match the \emph{same} node in $\T_2$
across all trees, but average RF distance allows a different node to match
on each tree.
The other differences are that \emph{average} RF distance
normalizes by sequence length, and is symmetrized.
The RF distance between two trees
was defined by \citet{robinson1981comparison}
% JK - change edge->branch here to avoid confusion with tskit/ARG edge
to be the minimum number of branch contraction/expansion operations needed to move
from one tree to the other
(which they then show is equal to the number of edges 
that induce different splits on the labels).
A similar metric on ARGs could be defined
using the subgraph-prune-and-regraft moves used by \citet{deng2024robust}.

The dissimilarity $\dis(\T_1, \T_2)$ measures disagreement between \emph{topologies},
but not times.
If the ARG is dated~\citep{wohns2022unified}, 
we can naturally use the ``best match'' $\alpha$
to also compare times.
Empirically, dating error seems to be more or less homoskedastic on a log scale,
so we recommend 
using the weighted root-mean-squared error of log(times), computed as
\begin{align}\label{eqn:wrmse}
    \text{wRMSE}_t(\T_1, \T_2)
    = \sqrt{\frac{
        \sum_{n \in N_1} s(\T_1,n) \left(\log\left(1+ t^{(1)}_n\right) - \log\left(1+ t^{(2)}_{\alpha(n)}\right) \right)^2 
    }{
        \Vert{\T_1}\Vert
    } } ,
\end{align}
where the transformation is $t \mapsto \log(1 + t)$ to avoid $\log(0)$.
The mean is computed weighting by node span, so that a dating error
is more impactful for a node with a longer span.

The implementation of this method in \texttt{tscompare}
additionally produces relative values:
the ARG RF value (dissimilarity relative to $\T_1$)
and true proportion represented (similarity relative to $\T_2$).
The ARG RF (which we call ``ARF'') is defined to be the dissimilarity proportional
 to the total span of nodes in $\T_1$,
 \begin{align}\label{eqn:dissimilarity}
 	\operatorname{ARF}(\T_1,\T_2)=\frac{\dis(\T_1,\T_2)}{\Vert \T_1 \Vert}
 	=1-\frac{\similarity(\T_1,\T_2)}{\Vert \T_1\Vert} ,
\end{align}
and so if $\T_2$ represents the truth, is analogous to a false positive rate.
The \emph{true proportion represented} (TPR) is similarity between two trees relative 
to the total span of nodes in $\T_2$,
\begin{align}\label{eqn:true-proportion}
	\operatorname{TPR}(\T_1,\T_2) = \frac{\similarity(\T_1,\T_2)}
	{\Vert \T_2\Vert} ,
\end{align}
and is analogous to statistical power.
The outputs framed as proportions relative to one of the given tree sequences is more easily understood
for comparing pairs of tree sequences than the original dissimilarity and similarity, whose units are length of spans.
We compute dissimilarity and ARF of two tree sequences in \figref{fig:conceptual_discrepancy}.

\begin{figure}[!ht]
	\begin{center}
\includegraphics[height=1.5in, width=3in]{discrepancy_func_method_t1.pdf}
\includegraphics[height=1.5in, width=3in]{discrepancy_function_method_t2.pdf}
    \caption{
        For two tree sequences $T1$ and $T2$ the \emph{dissimilarity}, $\dis(T1,T2)$, matches nodes in $T1$ to nodes in $T2$
        based on identical sample sets.
        In this example, node $8$ has no match in $T2$ as there are no nodes in $T2$ with sample set $\left\{1,2,3\right\}$.
        Node $4$ has no match on $[4,5)$ and matches with node $9$ on $[5,6)$. Thus the maximal mapping for 4 should be to itself. On the rest of the genome, all of nodes match with their identical counterpart.
        This makes the dissimilarity $\dis(T1,T2)=3$ and $\operatorname{ARF}(T1,T2)=\frac{3}{46}$.
        \label{fig:conceptual_discrepancy}
    }
	\end{center}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Simulations}

To do this, we simulate ARGs containing full haplotypes using \msprime{}
\citep{kelleher2016efficient,baumdicker2021efficient},
with the \texttt{coalescing\_segments\_only} option set to False.
Although \msprime{} simulates many events that do not create a coalescence in some marginal tree,
it only outputs information for nodes which contain a coalescence
(i.e., are the MRCA of some pair of samples at some point on the genome).
Furthermore, by default it only outputs those segments of the genome
on which there is a coalescence.
Said another way, by default all ancestral nodes in the ARG
output by \msprime{} are the MRCA of some pair of samples at every point in the genome
on which they are represented.
However, here we are interested in those segments of genome
on which the nodes are \emph{not} coalescent;
i.e., where they are unary in the marginal trees.
Setting \texttt{coalescing\_segments\_only} to False includes this information:
any ancestral segments for which these coalescent nodes
are ancestral to any samples -- so, the unary portions of their spans as well.
However, this includes more information than we want:
we hope to recover those portions of ancestral haplotypes
on which the nodes are unary, but adjacent to a region of the genome where the node is not unary.
For instance, if a lineage carrying an ancestral segment of genome that spans $[a, c)$
coalesces with another spanning $[b, d)$, with $a < b < c < d$,
then the resulting node is only coalescent on $[b, c)$ but we hope using this algorithm
to extend the node's span to $[a, b)$ and $[c, d)$
(on which the node is unary).
However, following this example, the first lineage might carry a segment $[x, y)$
that is disjoint from the segment $[a, d)$.
We call these segments ``isolated non-coalescent segments'';
they have also been called ``trapped unary spans'' \citep[by][]{wong2024general}.
Such isolated segments will not be recovered by our algorithm,
and would likely be unrecoverable by any other method.
So, after simulation, we first remove these isolated, non-coalescent segments.
To give an idea of what proportion of the full spans of ancestral nodes
these isolated non-coalescent segments represent,
a simulation of 1000 samples
with genome length $5\times 10^7$, recombination rate $10^{-8}$, and population size $10^4$
has about half the total span of all nodes in isolated, non-coalescent segments.
For more discussion of these segments, see \citet{baumdicker2021efficient}.

We used simulations of several scenarios.
To include the effects of heterogeneous recombination rate,
in some we used stdpopsim \citep{adrion2020stdpopsim} 
to simulate chromosome 1 of \textit{Canis familiaris}
using the CanFam3 genetic map from \citet{campbell2016}.
\emph{``Constant dog''} simulations simulated this chromosome in a 
population of (constant) size $10^4$.
\emph{``Expanding dog''} simulations were similar, but
used a discrete-time Wright--Fisher model to simulate
a small population of 100 that expanded to 1,000 individuals ten generations ago,
which then doubled every generation to reach 512,000 individuals.
All jobs for which runtime was recorded were executed on an Intel Xeon Gold 6148 processor.
 Additionally, we compare accuracy between sequences modified from a ``true'' tree sequence
 using our dissimilarity methods. 
 These tree sequences have a population of 1,000 and recombination rate of $10^{-8}$ with samples ranging from 10 to 1,000 and genome length ranging from $10^6$ to $5\times 10^7$.

\section{Results}

%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Tree sequence compression and computation}

In the simple example in \figref{fig:conceptual},
extending haplotypes replaces three edges
($0 \to 3$ and $3 \to 4$ on the left tree, and $0 \to 4$ on the right tree)
by two edges ($0 \to 3$ and $3 \to 4$ on both trees).
If all edge endpoints were unique, then we'd expect \emph{every} edge to be extendable
on one of its ends
(except those pendant to the root and some of those adjacent to chromosome ends),
leading to a reduction in number of edges by a third.
Experiments with an earlier version of the algorithm showed that
if we only extend haplotypes on ``paths of length 1'', 
then the hypothesized reduction of 1/3 is achieved for long sequences.
It is possible for \algorithmref{alg:extend} to add edges, as in \figref{fig:extending_diagram},
but we still expect the number of edges to decrease by more than 1/3.
Indeed,
\figref{fig:speed_and_edges} shows that \algorithmref{alg:extend} nearly cuts the number of edges
in half, as long as the sequence is long enough.

\begin{figure}
    \centering
    \includegraphics{benchmarks/one_pop_results_ratios}
    \caption{
        Ratio of
        \textbf{(A)} number of edges, and
        \textbf{(B)} runtime for computing Tajima's $D$;
        before and after extending haplotypes.
        For instance, extending haplotypes reduces number of edges by about 50\%
        and statistic computation runtime by about 20\%
        for long sequences.
        Horizontal axis shows sequence length;
        colors show numbers of samples;
        with lines showing averages across replicates.
        The original tree sequence was simulated with the ``expanding dog'' expanding population
        and subset to various sizes;
        see Methods for details. 
        Absolute values are shown in Supplementary \figref{fig:speed_and_edges_other}.
        \label{fig:speed_and_edges}
    }
\end{figure}

This reduction in edges can also lead to a reduction in computation time for
algorithms using the succinct tree sequence data structure.
Indeed, \figref{fig:speed_and_edges} shows that computation time is reduced by 10--20\%
for a typical statistic (here, Tajima's $D$),
computed in an efficient incremental manner along the genome as implemented in \tskit.
As described in \citet{ralph2020efficiently}, for these incremental algorithms
the addition or removal of an edge requires updates
to the state of the parent node and all nodes ancestral to it.
Extending haplotypes yields a tree sequence with fewer edge removals and insertions, 
and thus requires fewer traversals to the roots.

Supplementary \figref{fig:speed_and_edges_constant} shows these results are not specific
to the demographic scenario.
Supplementary Figures~\ref{fig:speed_and_edges_other}
and~\ref{fig:speed_and_edges_other_constant}
also show that our implementation of \algorithmref{alg:extend} is quite efficient,
running at chromosome scale in seconds to minutes for hundreds or thousands of samples,
or minutes to hours for tens of thousands of samples

%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Accuracy with true trees}

Our next task is to confirm that the haplotypes extended by \algorithmref{alg:extend}
are indeed correct -- i.e., that in addition to compression, we are also gaining information.
To do this, 
we simulate ARGs containing full haplotypes using \msprime{},
apply the simplification algorithm~\citep{kelleher2018efficient,wong2024general}
to reduce these so that there are no unary nodes
(i.e., any node present in a marginal tree is a coalescent node or a sample),
and then apply \algorithmref{alg:extend} to the result
(see Methods for more detail).
The method can potentially extend the spans of each node
(additional span over which the node will be unary);
and we can quantify how much of these extended spans were in the original ARG
(and thus correctly extended).

%$msprime.sim_ancestry(1000, sequence_length=5e7, recombination_rate=1e-8, population_size=1e4, coalescing_segments_only=False, random_seed=1)$
%From the above output we see that ratio of unary span bordering coalescing between 
%- ExtendPaths/Truth = 0.4917
%- ExtendEdges/Truth = 0.27828
%
%The trapped unary span of our ground truth is 49\% of the TS.
%
%Total unary span bordering coalescing: 255151996720.0 (51.0\%)
%Total trapped unary span: 245038231157.0 (49.0\%)
%Total coalescing span: 99950000000.0
%Total unary span bordering coalescing: 125466159245.0 (100.0\%)
%Total trapped unary span: 0.0 (0.0\%)
%Total coalescing span: 99950000000.0

As seen in \figref{fig:node-spans},
the vast majority of span added by extending haplotypes is correct.
In this example (which is typical),
99\% of all added span is correct;
95\% of nodes have no incorrectly added span;
and those incorrectly added spans are nearly always a small fraction of the original span.
The added information is significant:
the algorithm typically increases spans (i.e., lengths of ancestral haplotypes)
by around 50\%.

\begin{figure}[!bht]
	\includegraphics{newplots_wo_ee/figure-4.pdf}
	\caption{
        The effect of extending haplotypes on per-node spans
        in an ARG simulated with $10^4$ diploid samples, $10^4$ population, and 
        recombination rate of $10^{-8}$ on a sequence of length $10^8$.
        \textbf{(A)}
        Distribution of total amount of span added across nodes by \algorithmref{alg:extend};
        note the log scale on the $y$ axis.
        \textbf{(B)}
        Amount of incorrectly added span, plotted against total span, by node.
        95\% of nodes have no incorrect span; of the remainder,
        nearly all have less than 5\% incorrectly added;
        see Supplementary \figref{fig:incorrect_ratio}.
        Plots \textbf{(C)} and \textbf{(D)} show total spans per node,
        ordered by total span in the original ARG (which includes unary nodes).
        Dark blue dots in both show the spans in this original ARG.
        In \textbf{(C)}, lighter green dots show
        span after removing unary spans using \textit{simplify},
        while in \textbf{(D)},
        lighter red dots show span after extending haplotypes,
        i.e., applying \algorithmref{alg:extend} to the simplified ARG.
    }
	\label{fig:node-spans}
\end{figure}

These statistics are also reflected at the genomic scale, using measures of dissimilarity.
Line ``SE'' in \figref{fig:dissimilarity} shows total amounts of span removed by simplification
and re-inferred by \algorithmref{alg:extend} (correctly and incorrectly).
(Lines labeled with ``I'' involve re-inference of the ARG; discussed next.)
The top row shows the proportion of the given ARG that does not match the original (ARF),
showing that the total amount of mis-matching span
produced by extending haplotypes is very small ($\approx 1\%$).
(Simplification does not produce dissimilarity.)
The bottom row shows the proportion of the original ARG that is represented in the given ARG (TRP).
This shows that a large proportion of the ARG can be removed by simplification,
indicating that coalescent nodes are unary
over a substantial portion of their spans.
(Since the \texttt{simplify} operations removes these unary portions of haplotypes,
the ``simplified'' line (``S'') on the right of Figure~\ref{fig:dissimilarity} shows
the proportion of nodes' spans on which they are not unary.)
However, \algorithmref{alg:extend} can correctly replace most of these portions of haplotypes,
especially with larger sample sizes and sequence lengths
(the ``simplified--extended'' line; ``SE'').
For instance, with 1,000 samples and a $5 \times 10^7$bp genome,
haplotypes are unary on about half their spans (on average),
and extending haplotypes can infer more than half of this span from coalescent information only.

\begin{figure}[!hbt]
	\begin{center}
        \includegraphics[width=\textwidth]{figure5.pdf}
	\end{center}
    \caption{
        Accuracy and sensitivity of extended haplotypes
        across a range of sample sizes, on a sequence of length $5\times10^7$ (top)
        and sequence lengths, with 1,000 diploid samples (bottom).
        For each, a simulated ARG containing unary haplotype spans
        was (i) \emph{simplified}, removing the unary spans,
        and (ii) \emph{inferred}, using \tsinfer{} on genotypes;
        then each of these had its haplotypes \emph{extended}.
        The inferred, then simplified, ARG is also shown.
        \textbf{Top row:} ARF, the dissimilarity to the true ARG,
        as proportion of haplotypes that are not represented in the true ARG
        (\eqnref{eqn:dissimilarity}).
        \textbf{Bottom row:} TPR, agreement to the true ARG
        as proportion of the true ARG that is represented.
        }     
    \label{fig:dissimilarity}
\end{figure}

%%%%%%%%%%%%%%%%%%
\subsection{Inferred ARGs}

So far, we have demonstrated that there is potentially ample information
in the coalescent-only trees to extend haplotypes.
Does this work with \emph{inferred} ARGs?
As illustrated by~\cite{wong2024general}, there is a significant diversity
in the structures inferred by current methods,
and here we focus on \tsinfer{}~\cite{kelleher2019inferring} which 
infers ARGs containing unary nodes as a byproduct of its inference algorithm.
% Note this seems pretty weak and is sure to be dinged. We could at least 
% comment on what we might observe with other inference methods? Would we 
% recommend that people extend_edges() for their SINGER or Relate ARGs?
A comparison across other ARG inference methods is left for future work.
We simulated ARGs containing unary-spanning haplotypes (as above), then
re-inferred ARGs from the associated genotypes and performed various operations
on the results.
First, \figref{fig:dissimilarity} shows that \tsinfer{}
has a substantial portion of already--extended haplotypes:
comparing the ``inferred'' (``I''; dotted green) line to the ``inferred-simplified'' (``IS''; dashed green) line
we see that inclusion of these unary spans increases the amount of correctly inferred material
by around 20\% (bottom panels),
but that roughly 10\% of the unary spans in the inferred ARG are incorrect (top panels).
Furthermore, comparing to the inferred-extend(``IE''; red) line, we see that extending the tree sequence output with tsinfer adds relatively little span.
Applying \algorithmref{alg:extend} to the inferred-and-then-simplified ARG
(``ISE''; solid green line)
produces an ARG with both less correct and incorrect span.
Additionally, the ``IE'' and ``ISE'' ARGs 
contain approximately the same number of edges (difference of $\approx 1\%$).
It is also helpful to note that accuracy
(i.e., proportion of the true ARG that is inferred; bottom panels)
greatly increases with larger sample sizes,
possibly due to resolution of polytomies.
(Recall that due to computational constraints,
``correct'' and ``incorrect'' spans are determined here by \emph{exact} match
of subtending samples.)

%reporting proportions of correct and incorrect haplotype spans in Figure~\ref{fig:dissimilarity}.
%Main points: in \figref{fig:dissimilarity}
%(green line) tsinfer gets a lot right (above 60\% with 1000 samples; bottom-right plot)
%    and does a lot better with more samples;
%    also, gets 10\%--25\% wrong (left panels)
%    with longer sequence, larger proportion is wrong (bottom-left panel)
%(red dotted)
%    simplifying removes a lot of the wrong stuff
%    but also a lot of the right stuff
%(red solid)
%    similarly extending the simplified one adds a bunch of right and also wrong stuff
%    
%
%It's also interesting to  note that inference
%(see \figref{fig:inferred_edge_counts})
%(a) increases number of edges;
%(b) but inferred-simplifies have even more suggesting unary stuff is parsimonious;
%(c) but then extending reduces to below the original.

\section{Discussion}

We began this study with the observation that the simple transformation
of \figref{fig:extending_diagram} would reduce the number of edges
in the succinct tree sequence representation ARGs.
This is essentially a recombination-based parsimony argument,
and we have shown that this line of reasoning
leads to ARGs that are substantially more compact and faster to operate on,
and that contain more complete information about true ancestral relationships.
These extended ancestral haplotypes manifest as unary nodes in the marginal trees.
Although a number of ARG inference methods may be taking advantage of this information,
it is our impression that this source of information is not widely appreciated.
In fact, due to the field's focus on marginal trees rather than haplotypes,
we had to develop a haplotype-aware measure of (dis)agreement between ARGs in order to study the accuracy of the proposed algorithm.

There are good reasons to think that lengthening the spans of ancestral haplotypes
could lead to substantial gains in accuracy of ARG inference.
For instance, information about inferring the age of a particular mutation
derives almost entirely from constraints at nearby, linked sites.
Extending ancestral haplotypes from one site into neighboring regions
conceptually allows information from those marginal trees
to inform age inference at that site as well.

We have also explored the degree to which \tsinfer{} already makes use of this information,
and whether this algorithm can be used to improve inference.
The results do not provide a clear ordering:
for instance, although \tsinfer-produced ARGs
have a substantial portion of correctly inferred unary haplotypes,
removing these with simplification decreases both ARF
(i.e., proportion of ``wrong'' haplotypes)
and TPR (the proportion of the truth that is correctly inferred).
Extending haplotypes restores a large amount of this correctly inferred span,
but also introduces incorrect spans.
Further work is needed to determine how the balance of ``true and false positive rates''
affects downstream uses.
The efficient computational tools we have implemented (in \tskit{})
should facilitate this exploration.

\paragraph{Ignorance and omission in an ARG}
As motivation, we presented above a ``historical'' view of ARGs --
i.e., that each aspect of an inferred ARG is intended to represent
a portion of some particular historical genome
(for instance, the MRCA of some set of sampled genomes).
Furthermore, \figref[A\&B]{fig:conceptual} implicitly takes the position
that relationships \emph{not} depicted in an ARG are implied to not exist.
An alternative interpretation
of the ARG depicted in \figref[A\&B]{fig:conceptual} would be that we have no information as to 
how node 2 inherited from node 4 on the right-hand interval,
rather than saying that the line of transmission specifically did not pass
through node 3.
The ``simplification'' algorithm~\citep{kelleher2018efficient}
and the Hudson algorithm for coalescent
simulation~\citep{hudson1983properties,kelleher2016efficient}
each specifically discard information about any such ``non-coalescent'' portions of ancestral haplotypes;
so for ARGs produced by these algorithms, the correct interpretation is that
the omission of unary spans reflects a lack of knowledge.
In this paper, we have shown that, for the most part, this missing information can be imputed.

\paragraph{Metrics on ARGs}
The dissimilarity $\dis(\T_1,\T_2)$ is clearly not a metric in the mathematical sense
(i.e., symmetric, nonzero distance between distinct points, and satisfying the triangle inequality).
This is by design: in practice it is not possible to infer all aspects
of the true ancestry of a set of samples (i.e., all their genetic ancestors who ever lived),
and so we wanted to quantify
``How much of the true relationships does this ARG represent?''
However, it is worth noting that the symmetrized version
$d(\T_1,\T_2) = \dis(\T_1,\T_2) + \dis(\T_2, \T_1)$ is a metric.
To see this, first suppose that we have a bijection between the nodes of $\T_1$ and $\T_2$,
and view each ARG as a subset of the space $[0,L) \times N \times N$,
where $N$ is the shared set of nodes.
Then, dissimilarity is the Lebesgue measure of the relative difference of the two sets:
$|\T_1 \setminus \T_2|$,
and so the symmetrized version is the measure of the symmetric difference
$|\T_1\Delta\T_2|$,
which is well-known to be a metric \citep{rudin1976principles}. 
If the two ARGs have the same number of nodes,
we can consider all bijections between their nodes.
The symmetric difference between $\T_1$ and $\T_2$,
related through each bijection is a metric.
Then the minimum over all such metrics will still be a metric
since the minimum over a finite number of metrics is also a metric.
This also extends to two tree sequences with different numbers of nodes,
$|N_1|\neq|N_2|$, as we can take the minimum over all possible matchings $T_1\to\T_2$ and $\T_2\to\T_1$.

The RF distance \citep{robinson1981comparison}
essentially counts the number of differing branches between two trees;
the averaged RF distance~\citep{kuhner2015assessing}
averages this distance across marginal trees, weighted by span along the genome.
The method we present here for measuring dissimilarity between topologies of ARGs
is a straightforward generalization
that takes into account span along the genome of inferred ancestral haplotypes
(and separates the metric into two pieces as discussed in the previous paragraph).
However, the RF metric has many undesirable properties --
for instance, moving a single tip can result in a tree with maximum distance to the original --
and there is a substantial literature giving more robust generalizations
\citep[reviewed by][]{llabres2021generalized}.
Many of these generalizations \citep[e.g.,][]{bocker2013generalized}
relax the requirement that the match between subtended sample sets be exact,
and weight matches in some way by the size of the dissimilarity.
We considered such definitions as well, but kept to the simple case
for computational tractability --
the generalization of \citet{bocker2013generalized} is NP-hard to compute, even for a single tree.
In the ARG literature, \citet{zhang2023biobankscale}
defines a metric (called ``ARG total variation distance'') that includes branch lengths,
in a way similar to \citet{robinson1979comparison} and \citet{kuhner1994simulation};
however, it is still applied to ARGs as an average over marginal trees,
without enforcement of identity across haplotypes;
it would be useful to extend our dissimilarity to include branch lengths.

\paragraph{Parsimony}
Much of the early work on ARG inference aimed to extract as much information as possible
out of the small datasets of the time,
and so, roughly speaking, integrated over possible ARGs
with the goal of inferring higher-level parameters:
mostly, scaled mutation rate and recombination rate
\citep[e.g.,][]{hudson1985statistical,griffiths1996ancestral,kuhner2000maximum,stephens2000inference,fearnhead2001estimating}.
However, the space of possible ARGs for a given dataset is extremely large,
and other work aimed to identify the minimum number of recombinations
needed to explain a given dataset under the infinite alleles model of mutation
\citep[e.g.,][]{hein1990reconstructing,myers2003bounds,song2005constructing},
which turns out to be NP-complete \citep{wang2001perfect}.
For instance, \citet{minichiello2006mapping}
used a heuristic algorithm to produce ``plausible'' ARGs
(i.e., those that explained the data with few mutations and recombinations),
and searched for associations with traits in the resulting ensemble of ARGs.
Our approach for extending haplotypes follows the same logic,
that an ARG with fewer recombination events is more parsimonious,
and thus more likely.
For this reason, it will occasionally be wrong even if the trees are correct,
although in practice this source of error is likely much smaller than error in tree inference itself.

\paragraph{Inference methods}
Clearly, there is ample information in genomic data to infer the extent of ancestral haplotypes even where they are unary,
and this has the potential to significantly inform downstream methods, especially dating.
However, currently not all ARG inference methods return ARGs with unary nodes.
For instance, we have seen that tsinfer returns ARGs to which the method presented here adds little --
so, seems to be already taking advantage of this source of information.
To our knowledge, methods based on threading (sequentially adding or removing haplotypes) such as
SINGER \citep{deng2024robust} and ARGNeedle \citep{zhang2023biobankscale} return ARGs without any unary nodes,
although there is no conceptual issue with incorporating unary haplotypes into threading algorithms,
and at least SINGER contains information about unary nodes in its internal format.
On the other hand, locally building trees separately in each segment of genome as is done by Relate \cite{speidel2019method}
simply does not provide this sort of haplotype-based information (although it might be added using a method like ours).

\bibliography{references}

\appendix
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

\begin{figure}
    \includegraphics{benchmarks/one_pop_results_absolute_values}
    \includegraphics{benchmarks/one_pop_results_timing}
    \caption{
        \textbf{(Above:)}
        Absolute values for ratios shown in \figref{fig:speed_and_edges}:
        \textbf{(left)} numbers of edges; and
        \textbf{(right)} runtime.
        \textbf{(Below:)} runtime for the \texttt{extend\_haplotypes}
        implementation of \algorithmref{alg:extend} provided in the \tskit{} library.
        For other details, see \figref{fig:speed_and_edges}.
        \label{fig:speed_and_edges_other}
    }
\end{figure}

\begin{figure}
    \includegraphics{benchmarks/constant_pop_results_ratios}
    \caption{
        As in \figref{fig:speed_and_edges} except that
        the original tree sequence was simulated with an constant population
        of size $10^4$ (using the ``constant dog'' scenario);
        see Methods for details.
        Absolute values are shown in Supplementary \figref{fig:speed_and_edges_other_constant}.
        \label{fig:speed_and_edges_constant}
    }
\end{figure}

\begin{figure}
    \includegraphics{benchmarks/constant_pop_results_absolute_values}
    \includegraphics{benchmarks/constant_pop_results_timing}
    \caption{
        As in \figref{fig:speed_and_edges_other} except that
        the original tree sequence was simulated with an constant population
        of size $10^4$ (using the ``constant dog'' scenario).
        \label{fig:speed_and_edges_other_constant}
    }
\end{figure}

\begin{figure}[!hbt]
	\includegraphics{newplots_wo_ee/figure-4b-supplement}
	\caption{Distribution of incorrectly added span percentages
		for the simulated tree sequence in \figref{fig:node-spans}.}
	\label{fig:incorrect_ratio}
\end{figure}

\begin{table}[!hbt]
\begin{center}
\begin{tabular}{|c|c|c|c|c|}
	\hline
	Initial & Iteration 1 & Iteration 2 & Iteration 3 & Iteration 4 \\
	\hline
	\hline
	113463	&	76301	&	76059	&	76057	&	76057 \\
	\hline
	113266 & 76084	&	75835	&	75833	&	75833 \\
	\hline
	114489 & 76703 & 76436 & 76434 & 76434 \\
	\hline
	114086	& 76550	& 76294	& 76291	& 76291\\
	\hline
\end{tabular}
\caption{Number of edges for each iteration of \textit{extend haplotypes} applied to 
simulated tree sequences with sample size $10^4$ and length $5\times 10^7$. 
Each of the simulations terminated by the fourth iteration, however 99\% of the edges
are removed within the first iteration.}
\label{tab:edge-counts}
\end{center}
\end{table}

\end{document}